Running `Earth_obs_cleaning` on MCD19A2: Results from CONUS

### 2022-04-27 R/CONUS_AOD.Rmd
add targets for paper_conus output formats

---
title: "Running `Earth_obs_cleaning` on MCD19A2: Results from CONUS"
header-includes:
  - \usepackage{float}
  - \floatplacement{figure}{H}  #make every figure with caption = h, this was the fix
output: pdf_document
always_allow_html: true
bibliography: references.json
---

@@

```{r basics, echo=FALSE}
suppressPackageStartupMessages(
  {library(targets)
    library(Just.universal)
    library(data.table)
    library(DT)
    library(patchwork)
    library(ggplot2)
    library(sf)
    library(Just.universal)
    library(knitr)
    library(kableExtra)
    # library(flextable)
  })
# set_flextable_defaults(fonts_ignore=TRUE)
```

@@

```{r table_cv_results, echo=FALSE}
tar_load(cv_summary_tables)
result <- (cv_summary_tables$stat)[, !"loss"]
kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result))) %>% 
  kable_styling(latex_options=c("HOLD_position","scale_down")) %>%
  row_spec(0,bold=TRUE) %>% column_spec(2:12, width = "1cm")
```

@@

```{r strata_year, echo=FALSE, eval=TRUE}
result <- dt[, performance_aod(.SD), keyby = year(aer_date)]
kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result))) %>% 
  kable_styling(latex_options=c("HOLD_position","scale_down")) %>%
  row_spec(0,bold=TRUE) %>% column_spec(2:13, width = "1cm")
```

@@

```{r strata_month, echo=FALSE}
result <- dt[, performance_aod(.SD), keyby = month(aer_date)]
kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result))) %>% 
  kable_styling(latex_options=c("HOLD_position","scale_down")) %>%
  row_spec(0,bold=TRUE) %>% column_spec(2:13, width = "1cm")
```

@@

```{r strata_qa, echo=FALSE}
result <- dt[, performance_aod(.SD), by = qa_best]
kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result))) %>% 
  kable_styling(latex_options=c("HOLD_position","scale_down")) %>%
  row_spec(0,bold=TRUE) %>% column_spec(2:13, width = "1cm")
```

@@

```{r strata_overpasses, echo=FALSE, fig.pos="H"}
dt[, overpasses := .N, by = c("Site_Name", "aer_date")]
result <- dt[, performance_aod(.SD), keyby = overpasses]
kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result))) %>% 
  kable_styling(latex_options=c("HOLD_position","scale_down")) %>%
  row_spec(0,bold=TRUE) %>% column_spec(2:13, width = "1cm")

```

@@

# Contiguous US state boundaries
# will replace with a region defined by Kodi later
conus = get_conus()
```

@@

```{r strata_noaa_regions, echo=FALSE, fig.pos="H"}
# performance by NOAA climate regions
result <- dt[, performance_aod(.SD), keyby = state.noaa.region]
kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result))) %>% 
  kable_styling(latex_options=c("HOLD_position","scale_down")) %>%
  row_spec(0,bold=TRUE) %>% column_spec(2:13, width = "1cm")

## map by NOAA regions
# map_regions <- ggplot(conus) + geom_sf()
# print(map_regions)
```

@@

In decreasing order of variable importance (from Terra collocated cross-validation predictions)

```{r shap_rank, echo=FALSE}
# what do the SHAPs tell us?
tar_load(initial_cv_l2_terra_conus)
shaps <- initial_cv_l2_terra_conus$"shap_score"
# sort by importance
df_shap <- data.frame(meanabsshap = round(sort(sapply(abs(shaps), mean), decreasing = TRUE), 4))
kbl(df_shap, col.names = "Mean |SHAP|")
# df_shap$variable <- row.names(df_shap)
# ft_shap <- flextable(df_shap[, c(2, 1)])
# ft_shap <- set_header_labels(ft_shap, list(variable = "Variable", meanabsshap = "Mean Absolute SHAP"))
# ft_shap
```

>>>>

---
title: "Running `Earth_obs_cleaning` on MCD19A2: Results from CONUS"
header-includes:
  - \usepackage{float}
  - \floatplacement{figure}{H}  #make every figure with caption = h, this was the fix
  - \usepackage{helvet} \renewcommand\familydefault{\sfdefault}
output_format: 
  pdf_document: default
  html_document: default
always_allow_html: true
bibliography: references.json
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = "output", output_format = "all") })
---

@@

```{r basics, echo=FALSE}
suppressPackageStartupMessages(
  {library(targets)
    library(Just.universal)
    library(data.table)
    library(DT)
    library(patchwork)
    library(ggplot2)
    library(sf)
    library(Just.universal)
    library(knitr)
    library(kableExtra)
    # library(flextable)
  })
# set_flextable_defaults(fonts_ignore=TRUE)
source(here::here("R/functions.R"))
source(here::here("R/paper_functions.R"))
```

@@

```{r table_cv_results, echo=FALSE}
tar_load(cv_summary_tables)
result <- (cv_summary_tables$stat)[, !"loss"]
kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result)),
    align = "l") %>% 
  kable_styling(position = "left", 
                latex_options=c("HOLD_position","scale_down")) %>%
  row_spec(0,bold=TRUE) %>% column_spec(1:12, width = "1.1cm")
```

@@

```{r strata_year, echo=FALSE, eval=TRUE}
result <- dt[, performance_aod(.SD), keyby = year(aer_date)]
kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result)),
    align = "l") %>% 
  kable_styling(position = "left", 
                latex_options=c("HOLD_position","scale_down")) %>%
  row_spec(0,bold=TRUE) %>% column_spec(1:13, width = "1.1cm")
```

@@

```{r strata_month, echo=FALSE}
result <- dt[, performance_aod(.SD), keyby = month(aer_date)]
kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result)),
    align = "l") %>% 
  kable_styling(position = "left", 
                latex_options=c("HOLD_position","scale_down")) %>%
  row_spec(0,bold=TRUE) %>% column_spec(1:13, width = "1.1cm")
```

@@

```{r strata_qa, echo=FALSE}
result <- dt[, performance_aod(.SD), by = qa_best]
kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result)),
    align = "l") %>% 
  kable_styling(position = "left", 
                latex_options=c("HOLD_position","scale_down")) %>%
  row_spec(0,bold=TRUE) %>% column_spec(1:13, width = "1.1cm")
```

@@

```{r strata_overpasses, echo=FALSE, fig.pos="H"}
dt[, overpasses := .N, by = c("Site_Name", "aer_date")]
result <- dt[, performance_aod(.SD), keyby = overpasses]
kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result)),
    align = "l") %>% 
  kable_styling(position = "left", 
                latex_options=c("HOLD_position","scale_down")) %>%
  row_spec(0,bold=TRUE) %>% column_spec(2:13, width = "1.1cm")
```

@@

# Contiguous US state boundaries
# will replace with a region defined by Kodi later
conus = st_transform(crs = crs.us.atlas, get_conus())
```

@@

```{r strata_noaa_regions, echo=FALSE, fig.pos="H"}
# performance by NOAA climate regions
result <- dt[, performance_aod(.SD), keyby = state.noaa.region]
kbl(result, col.names = gsub(pattern = "\\.|_", " ", names(result)),
    align = "l") %>% 
  kable_styling(position = "left", 
                latex_options=c("HOLD_position","scale_down")) %>%
  row_spec(0,bold=TRUE) %>% column_spec(2:13, width = "1.1cm")
```

```{r noaa_region_plot, echo=FALSE, dev='png', dpi=300}
## map of performance by NOAA regions
conus$state.noaa.region <- factor(
   sapply(conus$STUSPS, USE.NAMES = F,
       function(sa) if (sa %in% c("AK", "HI")) NA else
           names(noaa.regions)[which(sapply(noaa.regions, function(r)
               sa %in% r))]),
   levels = names(noaa.regions))
conus <- merge(conus, dt[, performance_aod(.SD), keyby = state.noaa.region], by = "state.noaa.region")
conus_melt <- melt(as.data.table(conus[, c("geometry", "state.noaa.region", "RMSE_raw", "RMSE_pred")]),
                   id.vars = c("geometry", "state.noaa.region"))
levels(conus_melt$variable) <- c("MAIAC AOD", "corrected AOD")
map_region_rmse <- ggplot(conus_melt, aes(geometry = geometry)) + 
  geom_sf(aes(fill = value)) + 
  geom_sf(data = dt[, .(.N, geometry = geometry[1]), by = Site_Name], aes(geometry = geometry, size = N), alpha = 0.1) + 
  facet_wrap(~variable, ncol = 1) + 
  coord_sf(default_crs = st_crs(conus)) + 
  scale_size("colocated\nobservations", range = c(1, 3)) + 
  scale_fill_continuous("Regional\nRMSE") + 
  theme_minimal() + 
  theme(axis.text = element_blank(),
        panel.grid.major = element_blank())
map_region_rmse
```

@@

In decreasing order of variable importance (from Terra collocated cross-validation predictions)
```{r shap_rank, echo=FALSE}
# what do the SHAPs tell us?
tar_load(initial_cv_l2_terra_conus)
shaps <- initial_cv_l2_terra_conus$"shap_score"
# sort by importance
df_shap <- data.frame(meanabsshap = round(sort(sapply(abs(shaps), mean), decreasing = TRUE), 4))
kbl(df_shap, col.names = "Mean |SHAP|") %>% row_spec(0, bold=TRUE)
# df_shap$variable <- row.names(df_shap)
# ft_shap <- flextable(df_shap[, c(2, 1)])
# ft_shap <- set_header_labels(ft_shap, list(variable = "Variable", meanabsshap = "Mean Absolute SHAP"))
# ft_shap
```
